Testing Quasar Unification: I. Radiative Transfer in Clumpy Winds

### 2015-12-18 paper.tex
few edits

\begin{figure} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/sed.png}
\caption
{
Synthetic spectra at two viewing angles, 
this time in frequency space and including the optical band,
compared to the non-BAL SDSS quasar composite. The spectra are normalised to the flux at 
$2000$\AA, then an offset of 2 is applied per spectrum for clarity. 
}
\label{fig:sed}
\end{figure} %fullpage

@@

In figure~\ref{fig:sed}, we show an $F_{\nu}$ spectrum with broader waveband coverage
that includes the optical, showing that our synthetic spectra 
also exhibit \ha\ and \hb\ emission. 
In this panel, we include a low inclination and 
also a very high inclination 
spectrum, which looks underneath the wind cone. This model shows 
strong line emission with very similar widths and line ratios to the quasar composites,
although velocity projection effects mean the Balmer lines are double peaked, 
as seen  in `disc emitter' systems.
The increased line equivalent widths (EWs) are due to a weakened continuum at high inclination from wind attenuation and 
disc foreshortening / limb darkening. This effect also 
leads to a redder continuum slope, as seen in quasars, which is
due to Balmer continuum and Balmer and Fe~\textsc{ii} line emission.
The high inclination viewing angle is a physically motivated quasar model,
but does show that a model such as this can naturally reproduce quasar emission lines
if the emissivity of the wind is increased {\em with respect to the disc continuum}.
In addition, it neatly demonstrates how a stratified outflow can naturally
reproduce the range of ionization states seen in quasars.
% This is because we do not fully solve the C level populations,
% and thus the level populations do not accurately reflect those
% as a full macro-atom with a 
% full collisional rates between forbidden
% or semi-forbidden transitions, as would be required.
% The critical density of the C~\textsc{iii}]~1909\AA\, line 
% is $n_e\sim10^{9.5}~\rm{cm^{-3}}$ \citep{wei1988}, which is higher than much of the 
% outer portion of our wind. 
% We therefore expect a model with these 
% parameters to produce a C~\textsc{iii}]~1909\AA\ line
% with a proper treament. We note, however, the the presence of
% this line in the spectrum of quasars does {\em not} prohibit denser material having a significant contribution to the spectrum.

@@

Despite a number of successes, 
there are a some properties of the synthetic spectra
that are at odds with the observations. First, we find that the ratios of the 
EW of the \la\ and \mgline\ lines
to the EW of \civline\ is much lower than in the composite spectra. 
This problem has also been seen in simpler photoionization models for the 
BLR \citep{netzer1990}.
It may be that a larger region of very dense ($n_e\sim10^{10}$cm$^{-3}$) 
material is needed, which could correspond to a disc atmosphere or 
`transition region' (see e.g. Murray et al. 1995; Knigge et al. 1997). 
While modest changes to geometry may permit this, the initial grid search 
did not find a parameter space in which the \la\ or \mg\ EWs
were significantly higher (see section~\ref{sec:param_sens}. 
Second, we find that EWs increase
with inclination, to the extent that significantly denser
models can match the line EWs fairly well at low inclinations but will then
possess overly strong red wings to the BAL P-Cygni profiles at high inclinations
(see figures~\ref{fig:uvspec} and \ref{fig:sed}; also figure~\ref{fig:lobal}).
This angular dependence is expected from a model in which the BLR radiates
more isotropically than the disc, which in this case is foreshortened and limb darkened.
It appears to contradict observations, which show remarkably uniform emission
line properties in quasars and BALQSOs \citep{weymann1991,dipompeo2012b}. 
The angular distribution of the disc 
continuum and line emission
is clearly crucially important in determining the emergent broad 
line EWs, as suggested by, e.g., the analysis of \cite{risaliti2011}. 
We shall explore this question further in a future study.

>>>>

\begin{figure} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/sed.png}
\caption
{
Synthetic spectra at two viewing angles, 
this time in frequency space and including the optical band,
compared to the non-BAL SDSS quasar composite. The spectra are normalised to the flux at 
$2000$\AA, then an offset of 2 is applied per spectrum for clarity -- the dotted lines show the zero point of $F_\nu / F_{2000\AA}$ in each case.
}
\label{fig:sed}
\end{figure} %fullpage

@@

In figure~\ref{fig:sed}, we show an $F_{\nu}$ spectrum with broader waveband coverage
that includes the optical, showing that our synthetic spectra 
also exhibit \ha\ and \hb\ emission. 
In this panel, we include a low inclination and 
also a very high inclination 
spectrum, which looks underneath the wind cone. This model shows 
strong line emission with very similar widths and line ratios to the quasar composites,
although velocity projection effects mean the Balmer lines are double peaked, 
as seen  in `disc emitter' systems \citep[e.g.][]{eracleous1994}.
The increased line equivalent widths (EWs) are due to a weakened continuum at high inclination from wind attenuation and 
disc foreshortening / limb darkening. This effect also 
leads to a redder continuum slope, as seen in quasars, which is
due to Balmer continuum and Balmer and Fe~\textsc{ii} line emission.
The high inclination viewing angle is a physically motivated quasar model,
but does show that a model such as this can naturally reproduce quasar emission lines
if the emissivity of the wind is increased {\em with respect to the disc continuum}.
In addition, it neatly demonstrates how a stratified outflow can naturally
reproduce the range of ionization states seen in quasars.
% This is because we do not fully solve the C level populations,
% and thus the level populations do not accurately reflect those
% as a full macro-atom with a 
% full collisional rates between forbidden
% or semi-forbidden transitions, as would be required.
% The critical density of the C~\textsc{iii}]~1909\AA\, line 
% is $n_e\sim10^{9.5}~\rm{cm^{-3}}$ \citep{wei1988}, which is higher than much of the 
% outer portion of our wind. 
% We therefore expect a model with these 
% parameters to produce a C~\textsc{iii}]~1909\AA\ line
% with a proper treament. We note, however, the the presence of
% this line in the spectrum of quasars does {\em not} prohibit denser material having a significant contribution to the spectrum.

@@

Despite a number of successes, 
there are a some properties of the synthetic spectra
that are at odds with the observations. First, we find that the ratios of the 
EW of the \la\ and \mgline\ lines
to the EW of \civline\ is much lower than in the composite spectra. 
This problem has also been seen in simpler photoionization models for the 
BLR \citep{netzer1990}.
It may be that a larger region of very dense ($n_e\sim10^{10}$cm$^{-3}$) 
material is needed, which could correspond to a disc atmosphere or 
`transition region' (see e.g. Murray et al. 1995; Knigge et al. 1997).
\nocite{knigge1998} 
While modest changes to geometry may permit this, the initial grid search 
did not find a parameter space in which the \la\ or \mg\ EWs
were significantly higher (see section~\ref{sec:param_sens}. 
Second, we find that EWs increase
with inclination, to the extent that significantly denser
models can match the line EWs fairly well at low inclinations but will then
possess overly strong red wings to the BAL P-Cygni profiles at high inclinations
(see figures~\ref{fig:uvspec} and \ref{fig:sed}; also figure~\ref{fig:lobal}).
This angular dependence is expected from a model in which the BLR radiates
more isotropically than the disc, which in this case is foreshortened and limb darkened.
It appears to contradict observations, which show remarkably uniform emission
line properties in quasars and BALQSOs \citep{weymann1991,dipompeo2012b}. 
The angular distribution of the disc 
continuum and line emission
is clearly crucially important in determining the emergent broad 
line EWs, as suggested by, e.g., the analysis of \cite{risaliti2011}. 
We shall explore this question further in a future study.
